---
title: "Data 555 Interactive Widgets"
author: "Ying Xu"
date: '`r Sys.Date()`'
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
---

```{r}
# load package
pacman::p_load(here,readxl,haven,tidyverse,dplyr,gtsummary,ggplot2,plotly,DT)
# here-i-am
here::i_am("final_project.rmd") 
```

```{r cleaning}
# Import data
TKD_PaBS_2024 <- read_excel(here("raw_data", "Takunda PaBS Data 2024.xlsx"))
TTK_PaBS_2024 <- read_sav(here("raw_data", "data2024version1.sav"))

# TKD cleaning
TKD <- TKD_PaBS_2024 %>%
  mutate(hand_wash_location = 
           if_else(E16ADoyouhaveahandwashingstationatyourhomestead == "Yes", 1, 0)) %>%
  mutate(ICFY_diet_total = 
      if_else(A3RootsamptubersWhitepotatoeswhiteyamscassavapl_A == "Yes", 1, 0) +     #root
      if_else(A9Eggschickenturkeyfowlduck_A == "Yes", 1, 0) +                         #eggs
      if_else(A11LegumesnutsampseedsAnyfoodsmadefrombeanspeas == "Yes", 1, 0) +       #legumes nuts
      if_else(A1Isyourchildstillbreastfeeding == "Yes", 1, 0)) %>%                    #breast milk
  select(
    #demographics
    X02District, Z02Sexofrespondent, Z03Ageofrespondent, 
    Z04GenderedHouseholdType, Z05M2Male1529, Z05M3Male30, 
    Z05F2Female1529, Z05F3Female30,
    #exposure
    Z06TakundainterventionsparticipatedinWASH,
    #outcome
    hand_wash_location, ICFY_diet_total) %>%
  rename(District = X02District,
        Sex = Z02Sexofrespondent,
        Age = Z03Ageofrespondent,
        Gendered_HH_Type = Z04GenderedHouseholdType,
        Men_15_29 = Z05M2Male1529,
        Men_30 = Z05M3Male30,
        Female_15_29 = Z05F2Female1529,
        Female_30 = Z05F3Female30,
        Participate_WASH = Z06TakundainterventionsparticipatedinWASH) %>%
  mutate(Country = "Zimbabwe") %>%
  mutate(Sex = as.character(Sex)) %>%
  mutate(Participate_WASH = as.character(Participate_WASH))

TKD$Participate_WASH <- factor(TKD$Participate_WASH, levels = c(1, 0), 
                  labels = c("Yes", "No"))

# TTK cleaning
TTK <- TTK_PaBS_2024 %>%
  mutate(hand_wash_location = if_else(f06a %in% c(1, 2), 1, 0)) %>%
  mutate(ICFY_diet_total = f02b37 + 
           f02b32 + #root
           f02b35 + #eggs
           f02b33 + #legumes nuts
           f02b31   #breast milk
         ) %>%
  select(
    #demographics
    c02, c03, c04, c14, c15, c12, c13,
    #exposure
    c05.9,
    #outcome
    hand_wash_location, ICFY_diet_total
    ) %>%
  rename(Sex = c02,
        Age = c03,
        Gendered_HH_Type = c04,
        Men_15_29 = c14,
        Men_30 = c15,
        Female_15_29 = c12,
        Female_30 = c13,
        Participate_WASH = c05.9) %>%
  mutate(Country = "Malawi") %>%
  mutate(Sex = as.character(Sex)) %>%
  mutate(Participate_WASH = as.character(Participate_WASH))


TTK$Gendered_HH_Type <- factor(TTK$Gendered_HH_Type, levels = c(1, 2, 3, 4, 5),
                 labels = c("Adult Male no Adult Female (MnF)",
                            "Adult Female no Adult Male (FnM)",
                            "Adult Male and Adult Female (M&F)",
                            "Female Child no Adults (CNAF)",
                            "Male Child no Adults (CNAM)"))
TTK$Sex <- factor(TTK$Sex, levels = c(1, 2), 
                  labels = c("Male", "Female"))
TTK$Participate_WASH <- factor(TTK$Participate_WASH, levels = c(1, 2), 
                  labels = c("Yes", "No"))

# Combine dataset
combined <- bind_rows(TKD, TTK)
combined2 <- combined %>%
  filter(!is.na(ICFY_diet_total))
```

# Introduction

## Row {data-height="400"}

### Data Source

This is a Capstone Project, the data is retrieved from the Takunda and Titukulane programs, two USAID-funded resilience initiatives in Zimbabwe and Malawi, respectively, aim to achieve sustainable, equitable, and resilient food, nutrition, and income security through integrated interventions, including maternal and caregiver participation in Care Groups, farming schools, and income generating activities (IGAs). The data is collected from households who were sampled through intervention areas in Takunda and Titukulane that participated in the FY24 PaBS surveys.

### Dashboard Impact

This dashboard provides critical insights into the effectiveness of WASH interventions on children's dietary diversity, informing policymakers and stakeholders in public health. By clearly visualizing the link between WASH practices and dietary diversity, this dashboard helps program implementers identify gaps and areas for improvement. Its findings support evidence-based decisions, enhancing the impact of public health interventions.


# Interactive Boxplot

## Row {data-height="400"}

### Dietary Diversity by WASH Participation

KEY TAKE AWAYs: This section provides an analysis of ICFY, which is a measure of children dietary diversity scores in relation to participation in WASH programs. The larger the number the more diverse the dietary is. The hover tooltip provides interpretative summary statistics such as the median and interquartile range, allowing users to quickly assess differences between the groups. The interactive boxplot below illustrates the distribution of dietary diversity scores among different WASH participation groups.

```{r}
plot_ly(data = combined,
        x = ~Participate_WASH,
        y = ~ICFY_diet_total,
        type = 'box',
        boxpoints = FALSE,  # do not show individual data points
        hovertemplate = paste(
          "<b>WASH Participation:</b> %{x}<br>",
          "<b>Median ICFY Dietary Diversity:</b> %{median:.2f}<br>",
          "<b>Q1 (25th Percentile):</b> %{q1:.2f}<br>",
          "<b>Q3 (75th Percentile):</b> %{q3:.2f}<extra></extra>"
        )) %>%
  layout(xaxis = list(title = "WASH Program Participation", titlefont = list(size = 16)),
         yaxis = list(title = "ICFY Dietary Diversity (unit)", titlefont = list(size = 16)),
         width = 1100,
         height = 500)
```

# Data Table

## Row {data-height="400"}

### Participant Data

KEY TAKE AWAYs: This section offers interactive features that enhance data exploration and analysis. Users can efficiently locate specific entries using the search bar, sort columns in ascending or descending order by clicking on the column headers, and adjust the number of displayed rows per page. These functions facilitate a tailored and efficient examination of the dataset.

```{r}
# Interactive Data Table with Filtering and Sorting
datatable(combined2, options = list(
  pageLength = 10,
  autoWidth = TRUE,
  searchHighlight = TRUE,
  dom = 'Bfrtip',
  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
))
```

# Link to Github Repository

## Row {data-height="400"}

<https://github.com/Li1yxu/Data555_Final_Project>
